\documentclass{../../mathnotes}

\usepackage{tikz-cd}
\usepackage{todonotes}

\title{Introduction to Algebraic Topology PSET 4}
\author{Nilay Kumar}
\date{Last updated: \today}


\begin{document}

\maketitle

\begin{prop}
    Hatcher exercise 1.1.10
\end{prop}
\begin{proof}
    Consider two loops in $X\times Y$ based at $(x_0,y_0)$. The first loop is purely in the $X$
    direction, denoted by $f:I\to X\times\{y_0\}$, while the second is purely in the $Y$ direction,
    denoted by $g:I\to \{x_0\}\times Y$. We wish to find a homotopy between the loops
    $f\diamond g$ and $g\diamond f$. This can be done by continuously transporting the 
    basepoint of $f$ along $g$ via $h_t:I\times I\to X\times Y$ given by
    \[
        h_t(s)=
        \begin{cases}
            (x_0, g(2s)) & 0\leq s \leq \frac{t}{2}\\
            (f(2s-1), g(t)) & \frac{t}{2} \leq s \leq \frac{t+1}{2}\\
            (x_0, g(2s-1)) & \frac{t+1}{2} \leq s \leq 1
        \end{cases}.
        \]
    This homotopy is continuous by continuity of $f,g$ and it clearly takes
    $f\diamond g$ at $t=0$ to $g\diamond f$ at $t=1$.
\end{proof}

\begin{prop}
    Hatcher exercise 1.1.12
\end{prop}
\begin{proof}
    Any morphism $\phi_*:\pi_1(S^1)\to\pi^1(S^1)$ is simply a morphism $\phi_*:\Z\to\Z$, which
    is determined by the image of its generator. Then, if $n=\phi_*(1)$, the map $\phi$ that
    induces $\phi_*$ is obviously the continuous map $\phi:S^1\to S^1$ given by
    $\theta\mapsto n\theta$, as $\phi$ takes the $[\omega_1]$ to $[\omega_n]$, as required
    by $\phi_*$.
\end{proof}

\begin{prop}
    Hatcher exercise 1.1.16
\end{prop}
\begin{proof}
    Throughout this proof, we use freely Hatcher Proposition 1.17: if a space $X$ retracts onto
    a subspace $A$, then the homomorphism $i_*:\pi_1(A,x_0)\to\pi_1(X,x_0)$ induced by the inclusion
    $i:A\hookrightarrow X$ is injective. Further, if $A$ is a deformation retract of $X$, then
    $i_*$ is an isomorphism. We also use the fact that the functor $\pi_1$ takes
    products to products.
    \begin{enumerate}[(a)]
        \item Let $X=\R^3$ with $A$ any space homeomorphic to $S^1$. The existence of a
            retraction $r:X\to A$ would imply an injective morphism $\Z\hookrightarrow 0$ of
            the integers into the trivial group, which is absurd.
        \item Let $X=S^1\times D^2$ with $A$ its boundary torus $S^1\times S^1$. The existence
            of a retraction $r:X\to A$ would imply an injective morphism
            $\Z\times\Z\hookrightarrow\Z$ as $D^2$ is contractible. This is of course impossible:
            if $(1,0)\mapsto m$ for any $m,n\in\Z$ and $(0,1)\mapsto n$, we find that
            $n(1,0)$ maps to the same element that $m(0,1)$ maps to; there are no injective group
            morphisms $\Z\times\Z\to\Z$.
        \item Let $X=S^1\times D^2$ and $A$ be the circle as shown in Hatcher. The inclusion of
            $A$ into $X$ induces a map from $\Z$ to $\Z$ (injective by the retraction) that takes
            the generator to just the homotopy class of the knotted circle inside the solid torus.
            Of course, the knot can be homotoped away by pulling the ends through each other, and
            hence the map takes the generator to zero. But then the map must be the zero map, and
            hence not injective.
        \item Let $X=D^2\vee D^2$ with $A$ its boundary $S^1\vee S^1$. Of course,
            $\pi_1(D^2\vee D^2)$ is the trivial group: any loop in $D^2$ is contractible
            and since any loop in $D^2\vee D^2$ can be pinched off into the composition
            of two loops each entirely in a single copy of $D^2$, every loop in $D^2\vee D^2$
            is contractible as well. For usch a retraction $r:X\to A$ to exist would imply
            an injective morphism from $\pi_1(S^1\times S^1)$ to the trivial group, which is
            possible only if all loops in $S^1\times S^1$ are contractible. This is clearly
            not the case, as one can consider a loop in just one of the copies of $S^1$, and
            hence no such retraction exists.
        \item Let $X$ be a disk with two points on its boundary identified and $A$ its boundary
            $S^1\vee S^1$. It is clear that $X$ is homotopy equivalent to $S^1$, as it deformation
            retracts to the diameter connecting the two identified points. Hence such a retraction
            would imply an injective morphism $\pi_1(S^1\vee S^1)\hookrightarrow \Z$. But note
            that clearly $\pi_1(S^1\times S^1)=\Z\times\Z\hookrightarrow\pi_1(S^1\vee S^1)$
            and hence this is impossible, as it would imply an injective morphism of $\Z\times\Z$
            into $\Z$. Of course, this is easier to see via the fact that $\pi_1(S^1\vee S^1)=\Z*\Z$.
        \item Let $X$ be the M\"obius band and $A$ be its boundary circle. It is clear that $X$
            deformation retracts onto its central circle, and thus $\pi_1(X,x_0)=\Z$ and $\pi_1(A,x_0)=\Z$
            for $x_0\in A\subset X$. Drawing out the M\"obius band as a square with left and right
            sides anti-identified, it is clear that the inclusion of the homotopy class of the boundary
            circle into $\pi_1(X,x_0)$ yields twice the generator, as the boundary circle goes
            around the core circle twice. But this would imply that the retraction induces a map
            that takes twice the generator to the generator (as the existence of a retraction implies
            that the inclusion composed with retraction yields the identity), which is impossible
            from $\Z$ to $\Z$.
    \end{enumerate}
\end{proof}

\begin{prop}
    Hatcher exercise 1.2.4
\end{prop}
\begin{proof}
    Let $X\subset\R^3$ be the union of $n$ lines through the origin. It is clear that $\R^3-X$
    deformation retracts onto a $2n$-punctured $S^2$, where the punctures are antipodal.
    Note that in the case of $n=1$, we can deformation retract $S^1-\{N,S\}$ to a single equatorial
    $S^1$, as each puncture is contained by a circle. Similarly, for higher $n$, we can fix
    a basepoint on the sphere (that is not one of the punctured points) and draw loops about each
    punctured point, hence forming the wedge sum of $2n-1$ circles (as the last point one is
    automatically contained in the `outside' of the others). Hence the $2n$-punctured $S^2$
    can be deformation retracted onto the $2n-1$ wedge sum $S^1\vee\cdots\vee S^1$, and we find
    that $\pi_1(\R^3-X)=\Z*\cdots*\Z$, where the free product is taken $2n-1$ times.
\end{proof}

\begin{prop}
    Hatcher exercise 1.2.8
\end{prop}
\begin{proof}
    If we attach a circle of one of the tori to a circle of the other, the cell complex can
    be drawn as below. The generators are clearly $a,b,c$ but due to the two attached 2-cells,
    the loops $aba^{-1}b^{-1}$ and $aca^{-1}c^{-1}$ are contractible. Hence the fundamental
    group, via Hatcher, has a presentation
    \[\langle a,b,c\mid aba^{-1}b^{-1}, aca^{-1}c^{-1}\rangle.\]
    Note that this is simply the group $\Z\times (\Z*\Z)$, as $[a,b]=1$ and $[a,c]=1$.  
\end{proof}

\end{document}
